Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis

### 2017-11-29 paper.tex
A few minor changes throughout.

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009} and available online at http://www2.mmm.ucar.edu/people/bryan/cm1/.
The fully-compressible equations of motion were integrated on a 3000-km-wide, 35-km-deep, axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
Horizontal turbulence was parameterized using a Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, their Eq. 17), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
This formulation allows for realistically-large horizontal mixing lengths near the hurricane's inner core, consistent with the results of \cite{Bryan2012}, while not over-representing horizontal turbulence in convection at outer radii.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} microphysics scheme and radiative heating tendencies were computed every two minutes using the RRTMG longwave and shortwave schemes \citep{Iacono}.
A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
Above 19 km, where few TCI observations were available, the temperature profile was assumed to be isothermal.
Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, the water vapor mixing ratio was set to zero above 16 km.
The vortex described in \citeauthor{RotunnoEmanuel} (\citeyear{RotunnoEmanuel}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   EQ. 18 OF BRYAN AND ROTUNNO. ALSO USE EQ. 19
   \end{equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ approach zero; thus, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where \textit{g} is gravitational acceleration and $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in unsaturated environments.
For mathematical simplicity, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in tihs section.}.
Taking the time derivative of Eq.\ref{eq:n2dry} yields the static stability tendency:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acknowledgments
Start acknowledgments here.

>>>>

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009} and available online at http://www2.mmm.ucar.edu/people/bryan/cm1/.
The fully-compressible equations of motion were integrated on a 3000-km-wide, 35-km-deep, axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
Horizontal turbulence was parameterized using the Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, pg. 1773), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
This formulation allows for realistically-large horizontal mixing lengths near the hurricane's inner core, consistent with the results of \cite{Bryan2012}, while not over-representing horizontal turbulence in convection at outer radii.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} microphysics scheme and radiative heating tendencies were computed every two minutes using the RRTMG longwave and shortwave schemes \citep{Iacono}.
A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
Above 19 km, where few TCI observations were available, the temperature profile was assumed to be isothermal.
Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, relative humidity was set equal to 50\% above 11.5 km (the level above which temperature dropped below -40\textdegree C).
The vortex described in \citeauthor{RotunnoEmanuel} (\citeyear{RotunnoEmanuel}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   EQ. 18 OF BRYAN AND ROTUNNO. ALSO USE EQ. 19
   \end{equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ approach zero; thus, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where \textit{g} is gravitational acceleration and $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in unsaturated environments.
For mathematical simplicity, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in this section.}.
Taking the time derivative of Eq.\ref{eq:n2dry} yields the static stability tendency:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acknowledgments
George Bryan, Jeff Kepert, and Rob Fovell.
